``` {r setup, echo=FALSE, message=FALSE, include=FALSE, error=FALSE}
library(GenomicRanges, warn.conflicts=F)
library(magrittr)
library(parallel)
library(ggplot2)

setwd("/data/analysis_code/")
options(knitr.figure_dir = "figure2e_mRNA_and_pausing_index")
source("shared_code/knitr_common.r")
source("shared_code/granges_common.r")
source("shared_code/metapeak_common.r")
source("shared_code/sample_common.r")
source("shared_code/heatmap_common.r")
```

# Figure 2e mRNA level and Pausing index at genes with different paused Pol II half-life

**Author:** [Wanqing Shao](mailto:was@stowers.org)

**Generated:** `r format(Sys.time(), "%a %b %d %Y, %I:%M %p")`


## Overview 

Check mRNA level and pausing index at genes with different paused Pol II half-life. 

kc167 mRNA data were downloaded from flybase

Pausing index is defined as the average amount of Pol II per base-pair in a 41 bp window around pausing position, devide by the average amount of Pol II base-pair in a 101bp window 200 bp downstream of the pausing site.

### mRNA level

```{r rna_level}
kc_mrna <- read.table("/data/public_data/flybase/kc167_rna.txt",header=F, sep="\t")
kc_mrna <- kc_mrna[, c(2,  8)]
colnames(kc_mrna) <- c("fb_g_id", "RPKM")
half_life_df <- get(load("./rdata/half_life_df.RData"))

rna_df <- half_life_df %>% merge(., kc_mrna)

rna_pval <- wilcox.test(subset(rna_df, quantile == "q1")$RPKM, subset(rna_df, quantile == "q5")$RPKM)$p.value %>% 
            format(., scientific=T, digit=2) 

rna_plot <- ggplot(rna_df, aes(x=quantile, y=log(RPKM, base=2))) + 
            geom_boxplot(fill=c("#F0949A", "#FAD790", "#B7C593", "#81B6A9", "#6C97CA")) +  
            ggtitle(paste0("RNA level pval=", rna_pval)) +
            theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
                  panel.background = element_blank(), axis.line.x = element_line(colour = "black"), 
                  axis.line.y = element_line(colour = "black"))+
            xlab("Paused Pol II half-life") +
            ylab("RPKM (log2)")
rna_plot

```

### Pausing index

```{r pausing_index}
dmso_polii <- load_bigwig("dmso_control_polii_spikein",sample_format = "data")
tss <- get(load("rdata/dme_mrna_unique_tss.RData"))


### filterring out genes that are shorter than 300 bp
all_tx <- get(load("./rdata/dme_mrna.gr.RData"))
half_life_tx <- all_tx[all_tx$fb_t_id %in% half_life_df$fb_t_id]
#summary(half_life_tx$width)
### so the shortest gene is 458 bp

half_life_tx <- all_tx[all_tx$fb_t_id %in% half_life_df$fb_t_id]
half_life_tx_info <- data.frame(fb_t_id = half_life_tx$fb_t_id , width = width(half_life_tx))

fp_polii <- load_bigwig("fp_polii")
polii_mid <- resize(tss, 201, "center") %>% align_and_score_chipnexus_peaks(., fp_polii, "center")
polii_down <- resize(polii_mid, 201, "start") %>% resize(., 1, "end")



pausing_index <- ((resize(polii_mid, 41, "center") %>%
                 nexus_regionSums(., dmso_polii)) / 41 ) /
                 ((resize(polii_down, 101, "center") %>%
                 nexus_regionSums(., dmso_polii) )/101)
pausing_index <- data.frame(fb_t_id = polii_mid$fb_t_id, pausing_index=pausing_index) %>% 
                 merge(half_life_df, .)

pausing_index_pval <- wilcox.test(subset(pausing_index, quantile == "q1")$pausing_index, subset(pausing_index, quantile == "q5")$pausing_index)$p.value %>% 
                      format(., scientific=T, digit=2) 

pausing_index_plot <- ggplot(pausing_index, aes(x=quantile, y=log(pausing_index, base=2))) + 
                      geom_boxplot(fill=c("#F0949A", "#FAD790", "#B7C593", "#81B6A9", "#6C97CA")) +  
                      ggtitle(paste0("Pausing index pval=", pausing_index_pval)) +
                      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
                            panel.background = element_blank(), axis.line.x = element_line(colour = "black"), 
                            axis.line.y = element_line(colour = "black"))+
                      xlab("Paused Pol II half-life") +
                      ylab("Pausing index (log2)")
pausing_index_plot
```

```{r}
sessionInfo()
```